---
title: "exploratory_analysis"
author: "Fang Liu(fl2569)"
date: "10/10/2021"
output: github_document
---
```{r message=FALSE}
#tidyverse pacakge (dplyr, ggplot2, etc)
library(tidyverse)

#figure size options
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#set theme settings
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r load_data}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

#lubricate() package is helpful when dealing with dates
```

### `group_by`
```{r}
weather_df %>% 
  group_by(name, month)
```

## count some things
```{r}
x =
  weather_df %>% 
  group_by(name, month) %>%
  summarize(n_obs = n())

x
```

'table' does NOT produce a data frame!!! Not really useful
```{r use_table}
weather_df %>% 
  pull(name) %>% 
  table()
```

Or we can also use `tabyl()` to produce results that are data.frames and thus can be formatted!!
```{r tabyl}
x2 = 
  weather_df %>% 
  janitor::tabyl(month, name)
x2
```

## General summaries 

Let's count, take means, st. deviations, etc
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_max = sd(tmax, na.rm = TRUE)
  )
#or we can use drop.na() in the pipe
```

## Formatting table outputs
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  knitr::kable(digits =2)
```

## group `mutate`
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE), #mean_tmax is the 'group-wise mean' 
    centered_tmax = tmax - mean_tmax
  ) %>% 
ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
#add column specific values
```

#### Window functions
Window functions take n inputs and return n outputs, and the outputs depend on ALL the inputs!
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(desc(tmax)) #the `desc` ranks tmax in descending order
  ) %>% 
  filter(tmax_rank < 2)
```

**Lagged variables**  
Offsets are used to compare an observation to it’s previous value. This is useful to find the day-by-day change in max temperature within each station over the year:
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax, n = 1), #what is the lag information I have one day before (i.e., Yesterday)
    tmax_diff = tmax - lagged_tmax
  ) %>% 
  summarize(diff_sd = sd(tmax_diff, na.rm = TRUE)) #st.deviation of the daily difference
```

## Limitations
What if my "summary" is a linear model... well... it doesn't work! 
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(cor_tmin_tmax= cor(tmin, tmax, use = "complete"))

weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  lm(tmax ~ tmin, data = .)

#WRONG EXAMPLE
weather_df %>% 
  group_by(name) %>% 
  summarize(lm = lm(tmax - tmin))
```


























