---
title: "viz_part1"
author: "Fang Liu"
date: "10/5/2021"
output: github_document
---

```{r set_up}
library(tidyverse)
library(ggridges)

#to set global options for figures sizes
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```

load in a dataset that we'll use often 
```{r load_data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplot 
tmax vs. tmin
```{r create_scatter}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```
You can save ggplots
```{r save_plot}
ggp_tmax_tmin = weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin
```

## Fancier plots!
Add... colors/lines/extra stuff.
```{r fancy_mapping}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) +  #added in 3 different lines
  facet_grid(. ~ name) #i don't want to separate out the rows

#use color = NULL to "get rid" of a aesthetic setting??
#in ggplot: use aesthetic "globally" 
#in geom_point(): local changes/only apply to the points
```

Let's make one more scatter plot.
```{r temp_and_prec}
weather_df %>%
  ggplot(aes(x = date, y = tmax, size = prcp)) +
  geom_point(alpha = .3) +
  facet_grid(. ~ name) + 
  geom_smooth(se = FALSE)
```

## use data manipulation as part of this
You can do data manipulation before creating your plots 
```{r data_manipulation}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

## Stacking geoms

Which geoms do you want?

```{r without_using_points}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth()
```

```{r hex_plot}
weather_df %>%
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex()
```

## Univariate plots 

Looking at the distribution of a single variable.
#### Histograms
```{r histogram}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() + 
  facet_grid(. ~ name)
  
#ggplot invented before pipes were invented
```

#### Density plots
```{r density}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3)
```

#### Box plots
```{r boxplot}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

#### violin plots
```{r violin}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin()
```

What about ridges? Good for looking at the shape of distributions 
```{r ridge}
#Note: library(ggridges) is needed 
weather_df %>%
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85, alpha = 0.8)

```

## Embedding plots

```{r fig.width = 6, fig.height = 5}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```

```{r fig.width = 12, fig.height = 5}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```


