---
title: "viz_part2"
author: "Fang Liu"
date: "10/10/2021"
output: github_document
---


```{r setup, message=FALSE}
library(tidyverse)
library(patchwork)
```

Load the NOAA wether data.
```{r load_data, message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r simple_scatter, mesage = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

## Lables
```{r label_options, message=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

## Scales
Might want to control over the location and specification of tick marks on the X or Y axis. These can be manipulated in `scale_x_*` and `scale_y_*` where * depends on the type of variable mapped to the x and y aesthetics (i.e. continuous vs discrete).

```{r scaling_options, message=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15"))
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", #Y is squared
    position = "right")
```

```{r color_options}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(name = "Location", h = c(100, 300)) 

#name = "Location" --> the name of the plot legend is "Location"
```


Discrete = TRUE because the color aesthetic is mapped to a **discrete** variable. If we are using precipitation data (i.e., color = 'prcp'), then we can omit this argument to get a continuous color gradient.
```{r viridis_color}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_temp_plot
```

## Themes 
Themes are used to modify non-data elements of a plot – they don’t change mappings or how data are render, but control things like background color and location of the the legend. For example, to shift the position of the legend from right(default) to underneath:
```{r}
ggp_temp_plot + 
  theme(legend.position = "bottom")
```
**Quick tip**: `legend.position = "none"` will remove the legend. This is helpful when multiple plots use the same color scheme or when the legend is obnoxious for some other reason.

Note: default theme is gray.
```{r background_theme}
ggp_temp_plot + 
  theme_bw() + 
  theme(legend.position = "bottom")

ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

#### Learning Assessment
* Revisit the plot showing `tmax` against `date` for each location. Use labels, scale options, and theme changes to improve the readability of this plot.
```{r assessment}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = prcp), alpha = .75) + 
  labs(
    title = "Temperature plot",
    x = "Date",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```

## Other setting Options
Note: copy and paste this code chunk to the beginning of every new file!
```{r eval=FALSE}
#tidyverse pacakge (dplyr, ggplot2, etc)
library(tidyverse)

#figure size options
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#set theme settings
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data argument in geom_*
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

More realistically, it’s sometimes necessary to **overlay data summaries** on a plot of the complete data. Depending on the setting, one way to do this is to create a “summary” data frame and use that when adding a new geom to a ggplot based on the full data.

Sometimes you want to show two or three fundamentally different plots in the **SAME** graphic: you may want to juxtapose a scatter plot and a box plot, or show scatter plots illustrating relationships between different variables. In this case, a solution is to:

1. create each of the panels you want separately
2. Combine panels using tools in the `patchwork` package

```{r patchwork_example}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```
NOTE: `patchwork` is intended to help combine panels into a plot that couldn’t be created better using data tidying steps

## Data manipulation, reordering

Reorder by hand (fct_relevel):
```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

Reorders name according to tmax values in each name (fct_reorder):
```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

#### Example 1):
Suppose I wanted to create a three-panel plot showing densities for tmax and tmin within each location.
* facet panels across the name variable
* create separate densities for tmax and tmin in each panel

Solution: recognize that tmax and tmin are separate observation types of a shared temperature variable! Thus, we can tidy the weather_df and make the plot directly.
```{r}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```

#### Example 2): 
The code below imports and tidies the `PULSE` data, and creates a plot showing BDI score across visits.
* using `pivot_longer` to organize the BDI score and visit time variables
* organizing the visit time variable into a factor with an informative ordering.
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```

#### Example 3): 
Add some data tidying steps to view pup-level outcomes (post-natal day on which ears “work”, on which the pup can walk, etc) across values of dose category and treatment day.
```{r example_3}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
#x or rows = day_of_tx
#y or columns = outcome
```

**Message**: Yo should be proficient in data wrangling tools when conducting exploratory analyses!!! 
