---
title: "viz_part1"
author: "Fang Liu"
date: "10/5/2021"
output: github_document
---
Embedding plots in an R Markdown document can also take a while to get used because there are several things to adjust.
* Size of the figure **created** by R is controlled using two of the three chunk options `fig.width`, `fig.height`, and `fig.asp` 
* plots that are a little wider than they are tall: set options to fig.width = 6 and fig.asp = .6. 
* Second is the size of the figure **inserted** into your document, which is controlled using `out.width` or `out.height` (padding around the sides of figures -> set out.width = "90%")
```{r set_up, message = FALSE}
library(tidyverse)
library(ggridges)

#to set global options for figures sizes
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```

load in a data set that we'll use often.We’ll be working with **NOAA weather** data, which is downloaded using `rnoaa::meteo_pull_monitors` function:
```{r load_data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplot 
tmax vs. tmin

```{r scatter_plot, message=FALSE}
ggp_tmax_tmin = weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin
```

## Fancier plots!
Add... colors/lines/extra stuff.
```{r fancy_mapping, warning=FALSE}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) +  #added in 3 different lines
  facet_grid(. ~ name) #i don't want to separate out the rows but want to separate columns using name

#use color = NULL to "get rid" of a aesthetic setting??
#in ggplot: use aesthetic "globally" 
#in geom_point(): local changes/only apply to the points
```

Let's make one more scatter plot.
```{r temp_and_prec}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .3, aes(size = prcp)) +
  facet_grid(. ~ name) + 
  geom_smooth(se = FALSE)
```

### Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).
```{r data_manipulation}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Stacking geoms

Which geoms do you want?

```{r without_using_points, message=FALSE}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth()
```

```{r hex_plot, message = FALSE}
weather_df %>%
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex()

#Note: shape only applies to points
```

## Univariate plots 

Looking at the distribution of a single variable.

#### Histograms
```{r histogram}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() + 
  facet_grid(. ~ name)
  
#ggplot invented before pipes were invented so we have to use pluses (+) rather than %>% pipes
```

#### Density plots
```{r density}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3, color = "blue") +
  geom_rug() #shows the raw data?
```

#### Box plots
```{r boxplot}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

#### violin plots
```{r violin}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

#### Ridge plots
Ridge plots were the *trendiest* plot of 2017, and were a replacement for both boxplots and violin plots. 
* They’re implemented in the `ggridges` package
* good when there is a lot of categories in which the shape of the distribution matters
```{r ridge}
#Note: library(ggridges) is needed 
weather_df %>%
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85, alpha = 0.8)
```


### Learning Assessment: Make plots that compare precipitation across locations. 
```{r}
#density
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 

#ridge
ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

#boxplot
ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 

weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```

## Embedding plots
```{r fig.width = 6, fig.height = 5}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```
